Classifying Physical Movement Modalities From Accelerometer Data
========================================================

In this report, we use machine learning algorithms to predict the category of movement that predict the manner in which human did the exercise. based on sensor data[1].

## Obtain the data

```{r}
#download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv', 'pml-training.csv', method="wget")
pmlLabeled <- read.csv('pml-training.csv')

#download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv', 'pml-testing.csv', method="wget")
pmlUnlabeled <- read.csv('pml-testing.csv')
```

## Data Partition

We separate our labeled data into a train/test set(80%/20%).

```{r}
library(caret)
set.seed(1009)
pmlInTraining <- createDataPartition(pmlLabeled$classe, p = 0.8, list = FALSE)
pmlTraining <- pmlLabeled[pmlInTraining, ]
pmlTesting <- pmlLabeled[-pmlInTraining, ]
```

## Tidy Data

### Remove empty and NA
```{r}
removeEmptyNA <- function(col) { sum(!(is.na(col) | col=="")) }
```
```{r}
pmlColumnFillCounts <- sapply(pmlTraining, removeEmptyNA )
pmlFullColumns <- names(pmlColumnFillCounts[pmlColumnFillCounts == length(pmlTraining$classe)])
```

### Subset Data
Users action is not related to time, window and the name column is removed.

The label information is also removed. 

```{r}
exclude <- c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window", "classe")
pmlFullColumns <- pmlFullColumns[!pmlFullColumns %in% exclude]
```

```{r}
pmlTrainingTrimmed <- pmlTraining[, pmlFullColumns]
```

## Training

Train a Generalized Boosted Regression Model with Stochastic Gradient Boosting.
Using the defaults.
```{r}
pmlTrainControl <- trainControl(method="none")
pmlTuneGrid <- data.frame(interaction.depth = 4, n.trees = 200, shrinkage = .1)
pmlModel <- train(pmlTrainingTrimmed, pmlTraining$classe, method="gbm", trControl=pmlTrainControl, tuneGrid=pmlTuneGrid)
pmlModel
```

### Cross Validation

```{r}
pmlTestingLabelsPred <- predict(pmlModel, newdata=pmlTesting[, pmlFullColumns])
pmlConfusionMatrix <- confusionMatrix(pmlTestingLabelsPred, pmlTesting$classe)
pmlConfusionMatrix
```

### Out-of-sample error

```{r}
(1 - pmlConfusionMatrix$overall['Accuracy'])[[1]]
```

## Prediction

```{r}
pmlAnswers <- predict(pmlModel, newdata=pmlUnlabeled[, pmlFullColumns])
pmlAnswers
```

## Reference
[1] Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6. 

